---
title: "Assignment 4: Exploratory Data Analysis"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Sunday, 9/28 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot & interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project data folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
library(tidyverse)

data <- read_csv("data/cjs.csv")

```

### Describe your data

Data author:Mathurin Aché

Link:<https://www.kaggle.com/datasets/mathurinache/cjs-dataset?resource=download>

Description:effects of growth regulators, number and length of internodes formed on trimmed silver maple trees

### Preview your data

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}
data <- read_csv(here("data/cjs.csv"))

data <- data %>%
  clean_names()

view(data)
```

### Summary stats

Before you make plots, you must also:

-   Show basic summaries of your dataset, including:

    -   `count()` for categorical variables (to see category frequencies)

    -   `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

```{r}
data %>%
count(n)

data%>%
  group_by(n, tl)%>%
  summarize(tl=mean(tl, na.rm= TRUE), 
          internode_1=mean(internode_1, na.rm=TRUE))
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

## Part 2:

Create **at least 4 plots**, each of a different type. At least one of your four plots must use different variables than the others.

Each plot should have:

-   A title and labeled axes

-   A short interpretation (2-3 sentences) underneath

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can be plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}
ggplot(data = data, aes(x = internode_1, fill = tree)) +
  geom_histogram() +
  theme_bw() + 
  scale_fill_viridis_d(option="A") + 
  labs(title = "internode 1 of trees",
       subtitle = "the length of each the first internode of each tree",
       x = "internode 1",
       y = "number of trees")

```

Interpretation:

### Plot 2

```{r}
ggplot(data = data, aes(x = br, y = tl, fill = tree)) +
  geom_boxplot()

```

Interpretation:

### Plot 3

```{r}
ggplot(data = data, aes(x = internode_2, y = internode_3)) +
  geom_point()

```

Interpretation:

### Plot 4

```{r}
ggplot(data = data, aes(x = n, y = tr)) +
  geom_point()

```

Interpretation:

## Push to GitHub to Submit!

Assignment is due Sunday 9/28 at 11:59 pm.
